# Introduction{#intro}
First class, introduction to `R` and how to use it

## General R usage

Set up your directory

```{r}
setwd(".")
```


Install and load `tidyverse`

```{r,results="hide",message=F,warnings=FALSE}
install.packages("tidyverse")
library(tidyverse)
```
## Basics :

* R comments start with #
* Use text editor and beware of ”smart quotes”
* To get help about anything use `?`: `?seq` will show you the manual to  use the function `seq` (a function which generates sequences of numbers)
* To assign a value to a variable use the assignment operator `<-`

## Explore dataset:

We will use the mpg dataset which is load by default with `tidyverse`

To have more info about this dataset:
```{r}
?mpg 
```
To see the first lines of a dataset:
```{r}
 head(mpg) 
```

To see the _structure_ of the data:
```{r}
 str(mpg)
```
You can see from the output that this is a `data.frame`:  a rectangular structure with column and lines (`?data.frame` if you want to learn more `data.frame`). As you can see each column may store data of different types: here `int` _ie_ numbers and `chrs` _ie_ string of characters.


To get the list of column names:
```{r}
names(mpg)
```

To get some simple summary statistics of the dataset:

```{r}
summary(mpg)
```


## For the quickest

```{r,message=FALSE}
library(ggplot2)
library(ggthemes)
```


PLOT 1:
```{r,warnings=F}
iris %>%
    mutate(Species = 'ALL') %>%
    bind_rows(iris) %>%
    ggplot(aes(Petal.Length, Petal.Width, color = Species)) +
    geom_point() +
    geom_smooth() +
    xlab('Petal Length') +
    ylab('Petal Width') +
    facet_wrap(~Species, scales = 'free') +
    theme_bw()
```
PLOT 2:

```{r}
mtcars %>%
    rownames_to_column() %>%
    mutate(rowname = forcats::fct_reorder(rowname, mpg)) %>%
    ggplot(aes(rowname, mpg, label = rowname)) +
    geom_point() +
    geom_text(nudge_y = .3, hjust = 'left') +
    coord_flip() +
    ylab('Miles per gallon fuel consumption') +
    ylim(10, 40) +
    theme_classic() +
    theme(
          plot.title = element_text(hjust = 0, size = 16),
          axis.title.x = element_text(face = 'bold'),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank()
          )
```
PLOT 3:

```{r}
mtcars %>%
    ggplot(aes(mpg, qsec, size = disp, color = as.factor(am))) +
    geom_point() +
    scale_colour_discrete(name ="Gear",
    breaks=c(0, 1),
    labels=c("Manual", "Automatic")) +
    scale_size_continuous(name = 'Displacement') +
    xlab('Miles per gallon') +
    ylab('1/4 mile time') +
    theme_light()
```

PLOT 4:

```{r}
economics %>%
    ggplot(aes(date, uempmed)) +
    geom_line() +
    expand_limits(y = 0) +
    ggtitle('Median duration of unemployment [weeks]') +
    ylab('Median duration of unemployment [weeks]') +
    ggthemes::theme_economist_white() +
    theme(axis.title.x = element_blank())
```


 

