# Mapping and exploring real datasets: the case study of Obesity

Second class, read data, map and explore then



## CSV files: import and use them

In a `csv` file each cell is separated by a special character, which is usually a comma (Comma Separated Values), although other character can be used as well. In Europe, semi column (`;`) is often used to avoid confusion as the comma is used for separate unit with decimal.

The usual way to import a `csv` file called `mydata.csv` in R is by using `read.csv` function:
```{r,eval=F}
mydata=read.csv("mydata.csv") #read the csv file mydata.csv and store the result in an object called mydata
```

Feel free to use Excel or Libreoffice to format data. Usually, those program don't use `csv` format to save file you usually create with such software. To save a table you created using Excel or something similar, you may need to use the `Export to csv` option.

## The CDC data sets

### Download the data 
The website [Center for Disease Control and Prevention](cdc.gov) offers various dataset that you can use and explore. 

But before using those data set you will need to organize them in a format readable and usable by `R` and `ggplot`. For the purpose of this class we give you the dataset already formatted in the way we will need.
The file we need are:
* [`Obesity_2013.csv`](data/Obesity_2013.csv)
* [`Obesity_2004.csv`](data/Obesity_2004.csv)
* [`Income_2013.csv`](data/Income_2013.csv)
* [`Obesity_data_2013_county.csv`](data/Obesity_data_2013_county.csv)

You can download them by clicking to above's links.
Then put all those `csv` files in a R directory (advice: to better organize your the folder your are working in you can create another subdirectory `data` and store all your `csv` file within the directory)

### Import the data in R

To import those files as data frames in R: 

```{r}
obesity_map_2013 <- read.csv("data/Obesity_2013.csv")
obesity_map_2004 <- read.csv("data/Obesity_2004.csv")
obesity_cov_2013 <- read.csv("data/Obesity_data_2013_county.csv")
```

### Visualize the data
A quick way to visualise those data is to use the `choropleth` package
Install and use a  mapping package
The function we will use to create county `choropleth` maps is called
`county_choropleth`
Pass it as a data frame with one column named region and one column named value

```{r, eval=F}
install.packages(c("choroplethr", "choroplethrMaps"))
library(choroplethr)
library(choroplethrMaps)
```

Other more complex package to handle map can be found: 
* [](https://www.gislounge.com/mapping-county-demographic-data-in-r/)


```{r, eval =F}
county_choropleth(obesity_map_2013)
```

## Exploring Data

```{r}
obesity_cov_2013 <- read.csv("data/Obesity_data_2013_county.csv")
```



Warm-up: Show the structure, first 2 values etc.
Create scatterplot: Obesity versus Income
Fit trendiness through scatterplots
Explore correlations between variables

Structure of the obesity data

## First several rows of data frame

Last several rows of data frame

Last five rows

Plotting obesity data
* Create a scatter plot in ggplot: Income vs obesity in USA
* Check correlation between income and obesity

## Plot income vs. obesity
```{r}
ggplot(data = obesity_cov_2013) + geom_point(mapping = aes(x=
Income, y= Obes__aa))
```

Two underscores

Change default (y-)axis label:
```{r}
ggplot(data = obesity_cov_2013) + 
geom_point(mapping = aes(x = Income, y = Obes__aa)) + 
ylab("Obesity (%)")
```


## Use log-transformed income

Sometime to be able to better grasp the nature of your data you need to transform them. For data that cross different scales we often use a logarithmic transformation:

```{r}
ggplot(data = obesity_cov_2013) + geom_point(mapping = aes(x=Log_income, y= Obes__aa)) +
ylab("Obesity (%)") +
xlab("ln(income)")
```


### Add linear trend line with 95% CI
```{r}
ggplot(data = obesity_cov_2013, aes(x= Log_income, y= Obes__aa)) + geom_point() + geom_smooth(method=lm)
```

### Add loess trend line with 95% CI
```{r}
ggplot(data = obesity_cov_2013, aes(x= Log_income, y= Obes__aa)) + geom_point() + geom_smooth()
```

### Correlation income vs. obesity
```{r}
 cor.test(obesity_cov_2013$Income,obesity_cov_2013$Obes__aa)
```

Using income

```{r}
 cor.test(obesity_cov_2013$Log_income,obesity_cov_2013$Obes__aa)
```

Using log(income)

### Try a few other correlations or plots

```
County_pop: population of county
County_houses: households in county
Urban_fract: % urban population in county
Diab__aa: Diabetes %
Leis__aa: “leisure” statistic
lalowihalfshare: food deserts
```


Try a facet plot…?

### Facet plot
```{r}
ggplot(data = obesity_cov_2013, aes(x= Log_income, y= Obes__aa)) + geom_point() +
geom_smooth(method=lm)+ facet_wrap(~ State.1, nrow=10)
```

`bes__aa`

How can we limit number of states?

`Log_income`

```{r}
ggplot(data = obesity_cov_2013[1:286,], aes(x= Log_income, y=Obes__aa)) + geom_point() + geom_smooth(method=lm)+ facet_wrap(~State.1, nrow=3) + ylab("Obesity")
```


One way:
plot only a range of rows


Facet plot

Two
underscores

One way:
Subset the data

```{r}
Alabama <- obesity_cov_2013[ which(obesity_cov_2013$State.1=='AL'),]
first12 <- obesity_cov_2013[ which(obesity_cov_2013$State <=12),]
```


Correlations within each state
(google it!)

You must translate this googled answer into your data

### Correlations within each state

```{r}
r <- by(obesity_cov_2013, obesity_cov_2013$State.1, FUN =
function(X) cor(X$Income, X$Obes__aa, method = "spearman"))
```


Now group by state and do the same thing, using `obesity_cov_2013`

```{r}
 func <- function(obesity_cov_2013) {return(data.frame(COR =
cor(obesity_cov_2013$Income, obesity_cov_2013$Obes__aa)))}
 ddply(obesity_cov_2013, .(State), func)
```

