# Big-Data Lab (January 30): More plotting practice and beginning with Twitter data



## Objectives of this lab

This tutorial will cover the following:
* Practice more plots in ggplot;
* Practice ggplots using the public health dataframe (obesity_cov_2013) from previous lab.
* Begin to use Twitter data – from an archived dataset
* If you want to access Twitter API yourself (not required), you’ll need your Twitter account (see below).

The R functions in this lab program should:
* Find the maximum number of pages to be queried
* Generate all the subpages that make up the reviews
* Scrape the information from each of them
* Combine the information into one comprehensive data frame





## R CODE FOR TWITTER ASSIGNMENT

PACKAGES
```{r}
library(ggplot2)
```


## Packages for Twitter exercise 
You will need to install some of them first 

```{r}
library(tidyverse)
library(lubridate)
```



## Exploring ggplot

```{r}
ggplot(data = diamonds) + geom_bar(mapping = aes(x=cut))

ggplot(data = diamonds) + geom_bar(mapping = aes(x=cut, color=cut))

ggplot(data = diamonds) + geom_bar(mapping = aes(x=cut, fill=cut))

ggplot(data = diamonds) + geom_bar(mapping = aes(x=cut, fill = clarity), position = "dodge")
# careful with parentheses above

ggplot(mpg, aes(class, hwy)) + geom_boxplot()

```



## MORE PRACTICE GGPLOT ON OBESITY AND COVARIATES 

Here  you will need to use data frame from obesity lab: `obesity_cov_2013` (you may have to read it again. Feel free to include results on your assignment.

```{r}
ggplot(data = obesity_cov_2013) + geom_boxplot(aes(State.1,Income))
ggplot(data = obesity_cov_2013) + geom_boxplot(aes(State.1,Obes__aa))
```

Replace x-axis label

```{r}
ggplot(data = obesity_cov_2013) + geom_boxplot(aes(State.1,Income))+xlab("State")
```

Or just remove x-axis label

```{r}
ggplot(data = obesity_cov_2013) + geom_boxplot(aes(State.1,Income))+xlab("")
```

Too many states on the on x-axis?  Try rotating the boxplot: with coord_flip()

```{r}
ggplot(data = obesity_cov_2013) + geom_boxplot(aes(State.1,Income))+coord_flip()
```

 Notice that the states are still addressed by xlab() when using coord_flip()

 Want average values for each state? 
Get mean value of columns 7 through 13 for each state with this command

```{r}
States_2013 <- aggregate(obesity_2013[, 7:11], list(State = obesity_2013$State.1), mean)
```

 `States_2013` now has  values averaged across counties (not population) for each state. Is it useful? Note that a boxplot of these state-averaged data shows much less!


## BEGINNING WITH TWITTER DATASET

Read data into dataframe:

```{r}
tweets <- read.csv("streamlined_tweets.csv")
```





